Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-04 paper/main.tex
Uupdates

\subsection{Using Large Language Models as Experts}
A lot of recent work has focused on finding pairwise causal direction among
variables using Large Language Models. In this analysis, instead of using the
oracle to decide the direction of edges, we used GPT4 to give us the direction
of the edges based on description of the variables. As GPT4 has already
memorized many of these models, we encoded the variable names using a random
string and provided the description of the variable to let it decide the
direction of the edge.

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
	\caption{The learned DAG}
\end{figure}

>>>>

\subsection{Using LLMs as Experts}
Recently, there has been a lot of work towards utilizing LLMs for causal
discovery including approaches ranging from determining pairwise edge
orientation \citep{Kiciman2023, Jin2024} to full causal discovery
\citep{Naik2023, Vashishtha2023}, and counterfactual reasoning
tasks\citep{Kiciman2023} (see \citet{Liu2024} for an overview) As our approach
utilizes expert knowledge for determining the orientation of the edges, this
opened up an opportunity to use an LLM for determining this. We tried applying
our procedure for causal discovery with using GEMINI as the expert. We used a
greedy approach where the pair of variables with the highest unexplained
correlation is oriented first. The prompt used to determine the orientation is
provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
description to determine the edge orientation. We applied this approach to the
adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.

\begin{figure}[t!]
	\centering
	\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert.}
	\label{fig:adult_llm}
\end{figure}
